---
title: "US Census Cancer Data"
output:
  html_document: default
  pdf_document: default
date: "2023-09-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Living since, nothing has been clear on what causes this malicious disease, cancer. Although doctors and researchers put effort into seeking the mystery, people suffered cancer and ultimately faced death. There are numerous causes leading to a ghastly disease, cancer: mainly lifestyle factors, inheritance, genetic disorders, environmental exposures and so on. Among these factors, the biggest factor is known as aging, reviewing data of which factors contribute to diverse cancer. It is pretty certain that aging and cancer are correlated, but in this data review, one is looking forward to evaluating how much (the figure, amount) they actually are. Doing so, anova test will be used through a programming app ® Studio. The advent of cancer is more frequent in the old generation than in the young, other statistics suggest. By figuring out specific data from this experiment, one aims to encourage people to do such actions; if it comes out they are actually correlated, one wishes the world acknowledges the alertness and encourages the elderly to take health screenings once a year. The notion that is distinct but many people don't take seriously is that severe outcome due to cancer is something that can be prevented by checking your body. This data analysis's objective is to figure out the correlation between aging and cancer, and whether they affect each other, deriving a negative consequence. And if so, one's underlying goal is to alert people they can actually stop your death.

# Including Plots

We will import a data from the internet (wondercdc).

```{r}
data<-read.table(
  "/Users/seo-yun/Downloads/United States and Puerto Rico Cancer Statistics, 1999-2019 Incidence.txt", 
  header=TRUE, sep = "\t")
```

install necessary packages.

```{r}
install.packages("Rcpp")
library(Rcpp)
install.packages("tidyverse")
library(tidyverse)
```

We are going to name the following data as 'data'; it got rid of 'Notes' and 'Year.Code'. -c is to delete something.

```{r}
data <- data %>% subset(select = -c(Notes, Year.Code))

colnames(data)
```

We are making clear that the data 'data' has the columns L.C.S until Count.

```{r}
data <- data %>% subset(select = c(Leading.Cancer.Sites, Leading.Cancer.Sites.Code, Year, Age.Groups.Code, Count))
View(data)
```

We are organising two columns: Year and sumCount. We made it clear that sumCount is the sum of the total counts. After shortening the columns to two, name it 'datas'.

```{r}
datas <- data %>% select(Count, Year) %>% group_by(Year) %>% summarise(sumCount=sum(Count))
View(datas)
```

We are plotting a graph with the following code (x=Year and y=sumCount). Set the title and change the y-axis name to Cancer Incidence to make it clear for people looking at it.

```{r}
library(ggplot2)
ggplot(datas,aes(x=Year, y=sumCount)) + geom_line() + 
  labs(title="Cancer Count Over Years", y = "Cancer Incidence")
```

# Only for Breast Cancer

We are going to pick up the Cancer incidence for only Breast Cancers. Name it 'datah'.

```{r}
datah <- data%>%filter(Leading.Cancer.Sites=="Breast") %>% group_by(Year) %>% summarise(sumCOunt=sum(Count))
View(datah)
```

With the data 'datah' we set above, we plot another graph and change the title to an adequate one.

```{r}
library(ggplot2)
ggplot(datah,aes(x=Year, y=sumCOunt)) + geom_line() + 
  labs(title="Breast Cancer Count Over Years", y = "Cancer Incidence")
```

# All Year Incidence Data

We are making a data which contains every age group who had cancer. Also make sure the sum of the counts' title is sumCOunt which is different from the former one which appeared earlier.

```{r}
AllYear <- data%>% group_by(Leading.Cancer.Sites,Leading.Cancer.Sites.Code,Age.Groups.Code,Count) %>%
  summarize(sumCOUnt=sum(Count))

View(AllYear)
```

We are ploting another graph but this time bar graph. 'dodge' means it is vertically shown and 'identity' means it will display it by stacking them up.

```{r}
ggplot(AllYear,aes(x=Age.Groups.Code, y=sumCOUnt)) + geom_bar(stat = "identity", position = "dodge") + 
  labs(title="Cancer Incidence Recent 20 Years by age group", x = "Age", y = "Cancer Incidence")
```

This is showing every cancer incidence for every age group by bar graph. However, it is very hard to identify which one is which as they are stacked up right next to each other.

```{r}
ggplot(AllYear,aes(x=Age.Groups.Code, y=sumCOUnt, fill=Leading.Cancer.Sites)) + geom_bar(stat = "identity", position = "dodge") + 
  labs(title="Cancer Incidence Recent 20 Years by age group", x = "Age", y = "Cancer Incidence")
```

Give unique names to each age group for a concise summary and look.

```{r}
unique(AllYear$Age.Groups.Code) ##unique/distinct info

AllYearDecade <- AllYear%>% 
  mutate(Age.Groups.Code = case_when(
    grepl("1",Age.Groups.Code)~"< 10 year old", 
    grepl("1-4",Age.Groups.Code)~"< 10 year old", 
    grepl("5-9",Age.Groups.Code)~"< 10 year old", 
    grepl("10-14",Age.Groups.Code)~"10s", 
    grepl("15-19",Age.Groups.Code)~"10s", 
    grepl("20-24",Age.Groups.Code)~"20s",
    grepl("25-29",Age.Groups.Code)~"20s",
    grepl("30-34",Age.Groups.Code)~"30s",
    grepl("35-39",Age.Groups.Code)~"30s",
    grepl("40-44",Age.Groups.Code)~"40s",
    grepl("45-49",Age.Groups.Code)~"40s",
    grepl("50-54",Age.Groups.Code)~"50s",
    grepl("55-59",Age.Groups.Code)~"50s",
    grepl("60-64",Age.Groups.Code)~"60s",
    grepl("65-69",Age.Groups.Code)~"60s",
    grepl("70-74",Age.Groups.Code)~"70s",
    grepl("75-79",Age.Groups.Code)~"70s",
    grepl("80-84",Age.Groups.Code)~"80+",
    grepl("85+",Age.Groups.Code)~"80+",
    TRUE ~ Age.Groups.Code
    ))
View(AllYearDecade)
```

Now there are less age groups (as 20-24 and 25-29 are categorised as 20s) the graph is looking much better. It is much more distinct.

```{r}
ggplot(AllYearDecade,aes(x=Age.Groups.Code, y=sumCOUnt, fill=Leading.Cancer.Sites)) + geom_bar(stat = "identity", position = "dodge") + 
  labs(title="Cancer Incidence Recent 20 Years by age group", x = "Age", y = "Cancer Incidence")
```

We are selecting only the top 5 cancer sites (top 5 frequent cancers). We are finding the sum of sumCOUnt. In conclusion, we are finding the total cancer incidence of the top 5 cancers.

```{r}
Top5Year<-AllYearDecade%>%
  group_by(Leading.Cancer.Sites)%>% 
  summarize(sumCOUNt=sum(sumCOUnt), .groups = "drop") %>%
  arrange(desc(sumCOUNt))%>%
  head(5)

View(Top5Year)
```

'Leading Cancer Sites' column only contains the top 5 cancers. The columns are also summarised as Leading Cancer Sites code, age groups, count, and sumCOUnt. 'to_graph_data' is another set of data.

```{r}
Top5Code<-factor(Top5Year$Leading.Cancer.Sites)
to_graph_data <- AllYearDecade %>%
  filter(Leading.Cancer.Sites %in% as.character(Top5Code)) %>%
  mutate(Leading.Cancer.Sites = factor(Leading.Cancer.Sites, levels = levels(Top5Code)))  
View(to_graph_data)
```

This is a bar graph which displays the top 5 Cancer Incidence depending on the age group (they all have different colours for better identification)

```{r}
ggplot(to_graph_data,aes(x=Age.Groups.Code, y=sumCOUnt, fill=Leading.Cancer.Sites)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(title="Top 5 Cancer Incidence Recent 20 Years by age group", 
       x = "Age", y = "Cancer Incidence", fill = "Leading.Cancer.Sites")

```
